Legacy Data Confounds Modern Genomics Studies

### 2018-08-17 legacy.tex
updated text for Aikens et al

Fine scale mutational signatures are observable on evolutionary scales by comparing the rates of mutations between populations.
In 2015, Harris reported 50\% more TCC ${\rightarrow}$ TTC mutations in European populations compared to African populations; this was replicated in a different cohort in 2017 \cite{Harris2015a}.
A more in depth analysis of these subtle mutational signatures was recently released in preprint by Aikens, Johnson and Voight.

This strong population enrichment of a mutational signature suggests important genetic or environmental differences in the history of each population \cite{Harris2015a}. 
Harris and Pritchard further identified distinct mutational spectra across a range of populations \cite{Harris2017a}. 
In particular, they identified a heterogeneous mutational signature within 1000 Genomes Japanese individuals. 
Such an axis of variation is intriguing because differences in mutational signatures accumulate over many generations. 
A systematic difference among the Japanese population would suggest sustained environmental or genetic differences across sub-populations within Japan with little to no gene flow.
We therefore decided to follow up on this observation, by using a newly sequenced dataset \sgcomment{description of cohort.}

@@

Harris and Pritchard reported an excess of *AC${\rightarrow}$*CC mutations in a portion of the Japanese individuals in the 1kGP.\cite{Harris2015a}
While trying to follow up on this observations in a larger and more recent Japanese cohort, we did not find this particular signature.
However, when comparing the allele frequencies between the Japanese individuals from the 1kGP and this larger dataset, we observed an unusually large number of private single nucleotide polymorphisms (SNPs), only found in one of the two groups.
This is unexpected given the similarity of the two populations, and suggests a technical artifact rather than a population structure effect. 
Moreover, these mismatches were maintained after filtering for low-quality regions of the human genome and standard metrics such as Hardy-Weinberg equilibrium.

@@

%\subsection{GWAS from other papers}
%The only way to distinguish some of these more covert false positives is to use statistical tests associating the quality metrics of each position relative to each individual.

>>>>

Fine scale mutational signatures are observable on evolutionary scales by comparing the rates of mutations between populations.
In 2015, Harris reported 50\% more TCC ${\rightarrow}$ TTC mutations in European populations compared to African populations; this was replicated in a different cohort in 2017 \cite{Harris2015a}.
A more in depth analysis of these subtle mutational signatures was recently released in preprint by Aikens, Johnson and Voight. \cite{}
In their study they expanded the investigations of mutational signals in the 1kGP by investigations to broader sequence contexts.

This strong population enrichment of a mutational signature suggests important genetic or environmental differences in the history of each population \cite{Harris2015a}. 
Harris and Pritchard further identified distinct mutational spectra across a range of populations \cite{Harris2017a}. 
In particular, they identified a heterogeneous mutational signature within 1000 Genomes Japanese individuals.
In addition, multiple 7-mer mutational signatures were also identified by Aikens et. al in the same populations. \cite{}
Such an axis of variation is intriguing because differences in mutational signatures accumulate over many generations.
A systematic difference among the Japanese population would suggest sustained environmental or genetic differences across sub-populations within Japan with little to no gene flow.
We therefore decided to follow up on this observation, by using a newly sequenced dataset \sgcomment{description of cohort.}

@@

Harris and Pritchard reported an excess of a 3-mer *AC${\rightarrow}$*CC mutational pattern in a portion of the Japanese individuals in the 1kGP.\cite{Harris2015a}
While trying to follow up on this observations in a larger and more recent Japanese cohort, we did not find this particular signature.
However, when comparing the allele frequencies between the Japanese individuals from the 1kGP and this larger dataset, we observed an unusually large number of private single nucleotide polymorphisms (SNPs), only found in one of the two groups.
This is unexpected given the similarity of the two populations, and suggests a technical artifact rather than a population structure effect. 
Moreover, these mismatches were maintained after filtering for low-quality regions of the human genome and standard metrics such as Hardy-Weinberg equilibrium.

@@

\subsection{Twins Research Human Genetics}
Genotype + Imputation + GWAS

\subsection{PLos One}
Using 1kGP data directly, estimating count of EBV
